Domain-Driven, Generative Data Model for BigPetStore

### 2014-07-23 bdcloud_paper/latex/paper.tex
small edits

\section{Introduction}
Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.  With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from realistic example applications.  For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.

BigPetStore is a big data application blueprint built around processing transaction data for a fictional chain of pet stores.  BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}. BigPetStore has been incorporated with the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application. One of BigPetStore's key features is the ability to scale from a single machine to a large cluster, making it easy to develop projects on a local machine and transfer the application to a large cluster for production testing.

To achieve the project's goals of providing high-quality, realistic examples, BigPetStore requires semantically-rich, complex data. At its core, BigPetStore relies on a generative data model for producing synthetic transaction data. Despite the growing number of real data sets now publically available, synthetic data generators have a number of advantages for BigPetStore over real data sets. The synthetic data generator can be packaged with BigPetStore, avoiding the cost, time, and infrastructure needed to host, transfer, and store large data sets. Synthetic data generators are scalable, allowing the user to choose how much data to generate -- a requirement for supporting BigPetStore's goal of running on both single machines and large clusters.  Licensing and privacy issues are avoided with the use of synthetic data sets. And lastly, synthetic data generators are tunable by the user, allowing the user to generate data with specific criteria amenable to testing.  For example, a user may generate transactions from a small number of purchasable items so that clustering results can be easily visualized.

@@

\subsection{Relational Data Model}
The model generates data for four types of records: stores, customers, transactions, and transaction items (Figure~\ref{fig:relational-data-model}).  Store records contain both a unique identifier and a location in the form of a 5-digit zip code. Customer records consist of a unique identifier, a name, and a location given as a 5-digit zip code. Transaction records consist of a customer identifier, a store identifier, and a time given in days since the beginning of the simulation. Transaction item records contain a transaction identifier and an item description.  The item description is a list of key-value pairs stored as a string.  Key-value pairs are used since item characteristics differ depending on the type of item.  For example, dog food has a brand, a flavor, a size (in lbs), and a per-unit cost, while poop bags have a brand, a color, a count, and a per-unit cost.

@@

The customer's zip code is chosen by sampling from the zip codes with the probabiltiy of each zip code $z$ proportionate to its weight $w_z$.

@@

The item category for each purchase is chosen by sampling from the item categories with the probability of each item category $i$ proportionate to its weight $w_i$. \textcolor{red}{$\lambda_i$ - maybe these should be per customer / item category instead of just per item category}

>>>>

\section{Introduction}
Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.  With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from the availability of realistic example applications.  For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.

BigPetStore is a big data application blueprint built around processing transaction data for a fictional chain of pet stores.  BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}. BigPetStore has been incorporated into the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application. One of BigPetStore's key features is the ability to scale from a single machine to a large cluster, making it easy to develop projects on a local machine and transfer the application to a large cluster for production testing.

To achieve the project's goals of providing high-quality, realistic examples, BigPetStore requires semantically-rich, complex data. At its core, BigPetStore relies on a generative data model for producing synthetic transaction data. Despite the growing number of real data sets now publically available, synthetic data generators have a number of advantages for BigPetStore over real data sets. The synthetic data generator can be packaged with BigPetStore, avoiding the cost, time, and infrastructure needed to host, transfer, and store large data sets. Synthetic data generators are scalable, allowing the user to choose how much data to generate -- a requirement for supporting BigPetStore's goal of running on both single machines and large clusters.  Licensing and privacy issues are avoided with the use of synthetic data sets. And lastly, the generated data can be customized by the user, allowing the user to generate data with specific criteria amenable to testing.  For example, a user may generate transactions from a small number of purchasable items so that clustering results can be easily visualized.

@@

\subsection{Description of Generated Data}
The model generates data for four types of records: stores, customers, transactions, and transaction items (Figure~\ref{fig:relational-data-model}).  Store records contain both a unique identifier and a location in the form of a 5-digit zip code. Customer records consist of a unique identifier, a name, and a location given as a 5-digit zip code. Transaction records consist of a customer identifier, a store identifier, and a time given in days since the beginning of the simulation. Transaction item records contain a transaction identifier and an item description.  The item description is a list of key-value pairs stored as a string.  Key-value pairs are used since item characteristics differ depending on the type of item.  For example, dog food has a brand, a flavor, a size (in lbs), and a per-unit cost, while poop bags have a brand, a color, a count, and a per-unit cost.

@@

The customer's zip code is chosen by sampling from the zip codes with the probabiltiy of choosing each zip code $z$ proportional to its weight $w_z$.

@@

The item category for each purchase is chosen by sampling from the item categories with the probability of choosing each item category $i$ proportional to its weight $w_i$. \textcolor{red}{$\lambda_i$ - maybe these should be per customer / item category instead of just per item category}
